Open source symbolic multibody dynamics software

### 2012-02-21 PetersonGedeHubbard2012.tex
Updated table and made a few edits to permutation section.

The complete configuration of the system is assumed to be described by
generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
All coordinates and speeds, not only the independent ones, are represented by
$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
relationships described in table \ref{table:assumptions}.  The system is
described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
independent.  This paper discusses the first order relationship between these
independent quantities ($n-l$ independent coordinates, $o-m$ independent
speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
coordinates and $o$ speeds.
%It is assumed that there are $l$
%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
%assumptions can be summarized by the following equations
%\begin{align}
%  \label{eq:configuration}
%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
%  \label{eq:velocity}
%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
%  \label{eq:acceleration}
%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
%  \label{eq:kindiffs}
%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
%  \label{eq:dyndiffs}
%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
%\end{align}
\begin{table}[htbp]
  \centering
  \caption{Constrained multibody system governing definitions and equations}

  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q}$ & $\mathbf{R}^n$ & Coordinates \\
    $\bm{\dot{q}}$ & $\mathbf{R}^n$ & Time derivatives of coordinates \\
    $\bm{u}$ & $\mathbf{R}^o$ & Speeds \\
    $\bm{\dot{u}}$ & $\mathbf{R}^o$ & Time derivatives of speeds \\
    $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{3}{c}{ } \\
    \multicolumn{2}{c}{Equation} & Description \\
    \hline
    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration (holonomic)
    constraints \\
    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity
    (nonholonomic) constraints \\
    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
    \times 1}$ & Acceleration constraints \\
    $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential
    equations \\
  \end{tabular}
  \label{table:assumptions}
\end{table}
The first three equations in table \ref{table:assumptions} represent
constraints derived purely from kinematic considerations at the configuration,
velocity, and acceleration levels, respectively.  The velocity constraints may
be nonholonomic or differentiated holonomic, while the acceleration constraints
are assumed to be differentiated versions of the velocity constraints.  The
fourth equation represents kinematic differential equations which relate
$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
The fifth equation represents the constraint free nonholonomic dynamic
differential equations~\cite{Kane1985}, and will necessarily by linear in
$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
in all other quantities.  This formalism is exceedingly generic; in practice,
the structure of the functions in table \ref{table:assumptions} can be
leveraged to simplify the following calculations.  This will be discussed
subsequently.

@@

Additionally, there are two requirements for this method to work, namely, both
$\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
must be nonsingular.  Because of the presence of the permutation matrices, this
implies proper selection of dependent coordinates and dependent speeds is
critical.  The permutation matrices effectively select the columns of the
constraint gradient matrices, and may be selected by examination of the
constraint gradient matrix through a SVD or other numerical techniques; for
small systems it may be possible to choose dependent speeds without resorting
to numerical techniques if the system trajectories stay with the configuration
space where these matrices are nonsingular.

>>>>

The complete configuration of the system is assumed to be described by
generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
All coordinates and speeds, not only the independent ones, are represented by
$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
relationships described in table \ref{table:assumptions}.  The system is
described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
independent.  This paper discusses the first order relationship between these
independent quantities ($n-l$ independent coordinates, $o-m$ independent
speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
coordinates and $o$ speeds.
%It is assumed that there are $l$
%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
%assumptions can be summarized by the following equations
%\begin{align}
%  \label{eq:configuration}
%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
%  \label{eq:velocity}
%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
%  \label{eq:acceleration}
%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
%  \label{eq:kindiffs}
%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
%  \label{eq:dyndiffs}
%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
%\end{align}
\begin{table}[htbp]
  \centering
  \caption{Constrained multibody system governing definitions and equations}
  \label{table:assumptions}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^n$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^o$ & Speeds and their time derivatives\\
    $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{3}{c}{ } \\
    \multicolumn{2}{c}{Equation} & Description \\
    \hline
    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration constraints \\
    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity constraints \\
    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
    \times 1}$ & Acceleration constraints \\
    $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential equations
  \end{tabular}
\end{table}
The first three equations in table \ref{table:assumptions} represent
constraints derived purely from kinematic considerations at the configuration,
velocity, and acceleration levels, respectively.  The velocity constraints may
be nonholonomic or differentiated holonomic, while the acceleration constraints
are assumed to be differentiated versions of the velocity constraints.  The
fourth equation represents kinematic differential equations which relate
$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
The fifth equation represents the constraint free nonholonomic dynamic
differential equations~\cite{Kane1985}, and will necessarily by linear in
$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
in all other quantities.  This formalism is exceedingly generic; in practice,
the structure of the functions in table \ref{table:assumptions} can be
leveraged to simplify the following calculations.  This will be discussed
subsequently.

@@

This method relies on the nonsingularity of $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$
and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$.  The implication of this is that
selection of dependent coordinates and dependent speeds is critical.  The
permutation matrices effectively select the columns of the constraint gradient
matrices, and may be selected by examination of the constraint gradient matrix
through a SVD or other numerical techniques; for small systems it may be
possible to choose dependent speeds without resorting to numerical techniques
if the system trajectories stay with the configuration space where these
matrices are nonsingular.
